### 2022-12-22 README.qmd
set up

---
format: gfm

execute: 
  freeze: auto
  eval: false
  echo: true
  output: false

jupyter: julia-1.8
---

# Endogenous Macrodynamics in Algorithmic Recourse

{{< include intro.qmd >}}