### 2022-12-22 sections/data_preprocessing/_synthetic_data.qmd
more work on transfer to this repo

---
title: Generating Synthetic Data
---

```{julia}
using Pkg; Pkg.activate("dev")
```

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
output_path = output_dir("synthetic")
www_path = www_dir("synthetic")
data_path = data_dir("synthetic")
```

```{julia}
using MLJ, DataFrames, CSV
n = 1000
p = 2

using Random
Random.seed!(42)

# Linearly separable:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.1)
df = DataFrame(X)
df.target .= ifelse.(y.==1,0,1)
CSV.write(joinpath(data_path, "linearly_separable.csv"),df)

# Overlapping:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.5)
df = DataFrame(X)
df.target .= ifelse.(y.==1,0,1)
CSV.write(joinpath(data_path, "overlapping.csv"),df)

# Circles:
X, y = make_circles(n; noise=0.15, factor=0.01)
df = DataFrame(X)
df.target = y
CSV.write(joinpath(data_path, "circles.csv"),df)

# Moon:
X, y = make_moons(n)
df = DataFrame(X)
df.target = y
CSV.write(joinpath(data_path, "moons.csv"),df)
```