### 2022-12-22 sections/data_preprocessing/_synthetic_data.qmd
progress

---
title: Generating Synthetic Data
---

```{julia}
#| echo: false

include("src/setup.jl")
eval(setup)
output_path = output_dir("synthetic") # output directory for artifacts
www_path = www_dir("synthetic")       # output directory for images
data_path = data_dir("synthetic")
```

```{julia}
n = 1000
p = 2

# Linearly separable:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.1)
df = DataFrame(X)
df.target .= ifelse.(y.==1,0,1)
CSV.write(joinpath(data_path, "linearly_separable.csv"),df)

# Overlapping:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.5)
df = DataFrame(X)
df.target .= ifelse.(y.==1,0,1)
CSV.write(joinpath(data_path, "overlapping.csv"),df)

# Circles:
X, y = make_circles(n; noise=0.15, factor=0.01)
df = DataFrame(X)
df.target = y
CSV.write(joinpath(data_path, "circles.csv"),df)

# Moon:
X, y = make_moons(n)
df = DataFrame(X)
df.target = y
CSV.write(joinpath(data_path, "moons.csv"),df)
```