### 2022-12-23 paper/sections/limitations.rmd

# Limitations and Future Work {#limit}

While we believe that this work constitutes a valuable starting point for addressing existing issues in Algorithmic Recourse from a fresh perspective, we are aware of several of its limitations. In the following, we highlight some of these limitations and point to avenues for future research.

## Private vs. External Costs

Perhaps the most crucial shortcoming of our work is that we merely point out that there exists a trade-off between private costs to the individual and external costs to the collective of stakeholders. We fall short of providing any definite answers as to how that trade-off may be resolved in practice. The mitigation strategies we have proposed here provide a good starting point, but they are ad-hoc, mechanical extensions of the existing AR framework. An interesting idea to explore in future work could be the potential for Pareto optimal Algorithmic Recourse, that is, a collective recourse outcome in which no single individual can be made better off, without making at least one other individual worse off. This type of work would be interdisciplinary and could help to formalize some of the concepts presented in this work.

## Experimental Setup

The experimental setup proposed here is designed to mimic a real-world recourse process in a simple fashion. In practice, models are in fact updated on a regular basis [@upadhyay2021robust]. We also find it plausible to assume that the implementation of recourse happens periodically for different individuals, rather that all at once at time $t=0$. That being said, our experimental design is a vast over-simplification of potential real-world scenarios. In practice, any endogenous shifts that may occur can be expected to be entangled with exogenous shifts of the nature investigated in Upadhyay et al. @upadhyay2021robust. We also make implicit assumptions about the utility functions of the involved agents that may well be too simple: individuals seeking recourse are assumed to always implement the proposed Counterfactual Explanations; conversely, the agent in charge of the model $M$ is assumed to always treat individuals that have implemented valid recourse as if they were truly now in the target class.

## Causal Modelling

In this work we have focused on popular counterfactual generators that do not incorporate any causal knowledge. The generated perturbations therefore may involve changes to variables that affect the outcome predicted by the black-box model, but not the true outcome. The implementation of such changes is typically described as **gaming** @miller2020strategic, although they need not be driven by adversarial intentions: in Example \@ref(exm:student), student applicants may dutifully focus on acquiring credentials that help them to be admitted to university, but ultimately not to improve their chances of success at completing their degree @barocas2017fairness. Preventing such actions may help to avoid the dynamics we have pointed to in this work. Future work would likely benefit from including recent approaches to AR that incorporate causal knowledge such as Karimi et al. @karimi2021algorithmic.

## Classifiers

For reasons stated earlier we have limited our analysis to differentiable linear and non-linear classifiers, in particular logistic regression and deep neural networks. While these sorts of classifiers have also typically been analyzed in the existing literature on Counterfactual Explanations and Algorithmic Recourse, they represent only a subset of popular machine learning models employed in practice. Despite the success and popularity of deep learning in the context of high-dimensional data such as image, audio and video, empirical evidence suggests that other models such as boosted decision trees may have an edge when it comes to lower-dimensional tabular datasets, such as the ones considered here (@borisov2021deep, @grinsztajn2022why).

## Data {#limit-data}

Largely in line with the existing literature on Algorithmic Recourse, we have limited our analysis of real-world data to three commonly used benchmark datasets that involve binary prediction tasks. Future work may benefit from including novel datasets or extending the analysis to multi-class or regression problems, the latter arguably representing the most common objective in Finance and Economics.