### 2022-12-23 sections/data_preprocessing/_synthetic_data.qmd

## Generating Synthetic Data

```{julia}
#| echo: false

include("src/setup.jl")
eval(setup)
output_path = output_dir("synthetic") # output directory for artifacts
www_path = www_dir("synthetic")       # output directory for images
data_path = data_dir("synthetic")
```

```{julia}
n = 1000
p = 2

# Linearly separable:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.1)
df = DataFrame(X)
df.target .= ifelse.(y.==1,0,1)
CSV.write(joinpath(data_path, "linearly_separable.csv"),df)

# Overlapping:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.5)
df = DataFrame(X)
df.target .= ifelse.(y.==1,0,1)
CSV.write(joinpath(data_path, "overlapping.csv"),df)

# Circles:
X, y = make_circles(n; noise=0.15, factor=0.01)
df = DataFrame(X)
df.target = y
CSV.write(joinpath(data_path, "circles.csv"),df)

# Moon:
X, y = make_moons(n)
df = DataFrame(X)
df.target = y
CSV.write(joinpath(data_path, "moons.csv"),df)
```

```{julia}
#| output: true

catalogue = load_synthetic()
function plot_data(data,title)
    plt = plot(title=uppercasefirst(replace(string(title),"_" => " ")))
    scatter!(data)
    return plt
end
plts = [plot_data(data,name) for (name, data) in catalogue]
plt = plot(plts..., layout=(1,4), size=(850,200))
savefig(plt, "paper/www/synthetic_data.png")
display(plt)
```

```{julia}
generate_artifacts(data_path)
```