### 2022-12-28 sections/experiments/_mitigation_strategies.qmd

## Mitigation Strategies

```{julia}
#| echo: false

include("src/setup.jl")
eval(setup)
output_path = output_dir("mitigation_strategies")
www_path = www_dir("mitigation_strategies")
```

```{julia}
models = [
    :LogisticRegression, 
    :FluxModel, 
    :FluxEnsemble,
]
opt = Flux.Descent(0.01) 
generators = Dict(
    :Generic=>GenericGenerator(opt = opt, decision_threshold=0.5),
    :Latent=>REVISEGenerator(opt = opt),
    :Generic_conservative=>GenericGenerator(opt = opt, decision_threshold=0.9),
    :Gravitational=>GravitationalGenerator(opt = opt),
    :ClapROAR=>ClapROARGenerator(opt = opt)
)
```

### Synthetic

```{julia}
max_obs = 1000
catalogue = load_synthetic(max_obs)
choices = [
    :linearly_separable, 
    :overlapping, 
    :circles, 
    :moons,
]
data_sets = filter(p -> p[1] in choices, catalogue)
```

```{julia}
experiments = set_up_experiments(data_sets,models,generators)
```

```{julia}
n_evals = 5
n_rounds = 50
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 5
T = 100
using Serialization
results = run_experiments(
    experiments;
    save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, T=T
)
Serialization.serialize(joinpath(output_path,"results_synthetic.jls"),results)
```

### Plots

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results_synthetic.jls"))
```

```{julia}
using Images
line_charts = Dict()
errorbar_charts = Dict()
for (data_name, res) in results
    plt = plot(res)
    Images.save(joinpath(www_path, "line_chart_$(data_name).png"), plt)
    line_charts[data_name] = plt
    plt = plot(res,maximum(res.output.n))
    Images.save(joinpath(www_path, "errorbar_chart_$(data_name).png"), plt)
    errorbar_charts[data_name] = plt
end
```

#### Line Charts

@fig-mit-line shows the evolution of the evaluation metrics over the course of the experiment.

```{julia}

#| fig-cap: "Line Charts"
#| fig-subcap: 
#|      - "California Housing"
#|      - "Circles"
#|      - "Credit Default"
#|      - "GMSC"
#|      - "Linearly Separable"
#|      - "Moons"
#|      - "Overlapping"
#| layout-ncol: 1
#| label: fig-mit-line
img_files = readdir(www_path)[contains.(readdir(www_path),"line_chart") .&& .!contains.(readdir(www_path),"latent")]
img_files = joinpath.(www_path,img_files)
for img in img_files
    display(load(img))
end
```

#### Error Bar Charts

@fig-mit-error shows the evaluation metrics at the end of the experiments.

```{julia}

#| fig-cap: "Error Bar Charts"
#| fig-subcap: 
#|      - "California Housing"
#|      - "Circles"
#|      - "Credit Default"
#|      - "GMSC"
#|      - "Linearly Separable"
#|      - "Moons"
#|      - "Overlapping"
#| layout-ncol: 1
#| label: fig-mit-error
img_files = readdir(www_path)[contains.(readdir(www_path),"errorbar_chart") .&& .!contains.(readdir(www_path),"latent")]
img_files = joinpath.(www_path,img_files)
for img in img_files
    display(load(img))
end
```

### Bootstrap

```{julia}
n_bootstrap = 100
df = run_bootstrap(results, n_bootstrap; filename=joinpath(output_path,"bootstrap_synthetic.csv"))
```

### Chart in paper

@fig-mit-paper shows the chart that went into the paper.

```{julia}
using DataFrames, Statistics
df = results[:overlapping].output
df = df[df.n .== maximum(df.n),:]
gdf = groupby(df, [:generator, :model, :n, :name, :scope])
df_plot = combine(gdf, :value => (x -> [(mean(x),mean(x)+std(x),mean(x)-std(x))]) => [:mean, :ymax, :ymin])
df_plot = df_plot[[name in [:mmd, :mmd_grid, :model_performance] for name in df_plot.name],:]
df_plot = df_plot[.!(df_plot.name.==:mmd .&& df_plot.scope.==:model),:]
df_plot = mapcols(x -> typeof(x) == Vector{Symbol} ? string.(x) : x, df_plot)
transform!(df_plot, :name => (X -> [x=="mmd" ? "MMD (domain)" : x for x in X]) => :name)
transform!(df_plot, :name => (X -> [x=="mmd_grid" ? "MMD (model)" : x for x in X]) => :name)
transform!(df_plot, :name => (X -> [x=="model_performance" ? "Performance" : x for x in X]) => :name)
transform!(df_plot, :generator => (X -> [x=="Generic" ? "Generic (γ=0.5)" : x for x in X]) => :generator)
transform!(df_plot, :generator => (X -> [x=="Generic_conservative" ? "Generic (γ=0.9)" : x for x in X]) => :generator)
transform!(df_plot, :model => (X -> [x=="FluxEnsemble" ? "Deep Ensemble" : x for x in X]) => :model)
transform!(df_plot, :model => (X -> [x=="FluxModel" ? "MLP" : x for x in X]) => :model)
transform!(df_plot, :model => (X -> [x=="LogisticRegression" ? "Linear" : x for x in X]) => :model)

ncol = length(unique(df_plot.model))
nrow = length(unique(df_plot.name))

using RCall
scale_ = 2.0
R"""
library(data.table)
df_plot <- data.table($df_plot)
model_order <- c("Linear", "MLP", "Deep Ensemble")
df_plot[,model:=factor(model, levels=model_order)]
library(ggplot2)
plt <- ggplot($df_plot) +
    geom_bar(aes(x=n, y=mean, fill=generator), stat="identity", alpha=0.5, position="dodge") +
    geom_pointrange( aes(x=n, y=mean, ymin=ymin, ymax=ymax, colour=generator), alpha=0.9, position=position_dodge(width=0.9), size=0.5) +
    facet_grid(
        rows = vars(name),
        cols =  vars(model), 
        scales = "free_y"
    ) +
    labs(y = "Value") + 
    scale_fill_discrete(name="Generator:") +
    scale_colour_discrete(name="Generator:") +
    theme(
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="bottom"
    ) +
    guides(fill=guide_legend(ncol=3))
temp_path <- file.path(tempdir(), "plot.png")
ggsave(temp_path,width=$ncol * $scale_,height=$nrow * $scale_ * 0.8) 
"""

img = Images.load(rcopy(R"temp_path"))
Images.save(joinpath(www_path,"paper_synthetic_results.png"), img)
```

```{julia}
#| label: fig-mit-paper
#| fig-cap: "Chart in paper"

Images.load(joinpath(www_path,"paper_synthetic_results.png"))
```

### Latent Space Search

```{julia}
generators = Dict(
    :Latent=>GenericGenerator(opt = opt, decision_threshold=0.5),
    :Latent_conservative=>GenericGenerator(opt = opt, decision_threshold=0.9),
    :Gravitational=>GravitationalGenerator(opt = opt),
    :ClapROAR=>ClapROARGenerator(opt = opt)
)
```

```{julia}
experiments = set_up_experiments(data_sets,models,generators)
```

```{julia}
n_evals = 5
n_rounds = 50
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 5
T = 100
using Serialization
results = run_experiments(
    experiments;
    save_path=output_path, evaluate_every=evaluate_every, n_rounds=n_rounds, n_folds=n_folds, T=T, 
    latent_space = true
)
Serialization.serialize(joinpath(output_path,"results_synthetic_latent.jls"),results)
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results_synthetic_latent.jls"))
```

```{julia}
using Images
line_charts = Dict()
errorbar_charts = Dict()
for (data_name, res) in results
    plt = plot(res)
    Images.save(joinpath(www_path, "line_chart_latent_$(data_name).png"), plt)
    line_charts[data_name] = plt
    plt = plot(res,maximum(res.output.n))
    Images.save(joinpath(www_path, "errorbar_chart_latent_$(data_name).png"), plt)
    errorbar_charts[data_name] = plt
end
```

### Plots

#### Line Charts

@fig-mit-line-latent shows the evolution of the evaluation metrics over the course of the experiment.

```{julia}

#| fig-cap: "Line Charts"
#| fig-subcap: 
#|      - "Circles"
#|      - "Linearly Separable"
#|      - "Moons"
#|      - "Overlapping"
#| layout-ncol: 1
#| label: fig-mit-line-latent
img_files = readdir(www_path)[contains.(readdir(www_path),"line_chart") .&& contains.(readdir(www_path),"latent")]
img_files = joinpath.(www_path,img_files)
for img in img_files
    display(load(img))
end
```

#### Error Bar Charts

@fig-mit-error-latent shows the evaluation metrics at the end of the experiments.

```{julia}

#| fig-cap: "Error Bar Charts"
#| fig-subcap: 
#|      - "Circles"
#|      - "Linearly Separable"
#|      - "Moons"
#|      - "Overlapping"
#| layout-ncol: 1
#| label: fig-mit-error-latent
img_files = readdir(www_path)[contains.(readdir(www_path),"errorbar_chart") .&& contains.(readdir(www_path),"latent")]
img_files = joinpath.(www_path,img_files)
for img in img_files
    display(load(img))
end
```

### Bootstrap

```{julia}
n_bootstrap = 100
df = run_bootstrap(results, n_bootstrap; filename=joinpath(output_path,"bootstrap_latent.csv"))
```

### Chart in paper

@fig-mit-latent-paper shows the chart that went into the paper.

```{julia}
using DataFrames, Statistics
df = results[:overlapping].output
df = df[df.n .== maximum(df.n),:]
gdf = groupby(df, [:generator, :model, :n, :name, :scope])
df_plot = combine(gdf, :value => (x -> [(mean(x),mean(x)+std(x),mean(x)-std(x))]) => [:mean, :ymax, :ymin])
df_plot = df_plot[[name in [:mmd, :mmd_grid, :model_performance] for name in df_plot.name],:]
df_plot = df_plot[.!(df_plot.name.==:mmd .&& df_plot.scope.==:model),:]
df_plot = mapcols(x -> typeof(x) == Vector{Symbol} ? string.(x) : x, df_plot)
transform!(df_plot, :name => (X -> [x=="mmd" ? "MMD (domain)" : x for x in X]) => :name)
transform!(df_plot, :name => (X -> [x=="mmd_grid" ? "MMD (model)" : x for x in X]) => :name)
transform!(df_plot, :name => (X -> [x=="model_performance" ? "Performance" : x for x in X]) => :name)
transform!(df_plot, :generator => (X -> [x=="Latent" ? "Latent (γ=0.5)" : x for x in X]) => :generator)
transform!(df_plot, :generator => (X -> [x=="Latent_conservative" ? "Latent (γ=0.9)" : x for x in X]) => :generator)
transform!(df_plot, :model => (X -> [x=="FluxEnsemble" ? "Deep Ensemble" : x for x in X]) => :model)
transform!(df_plot, :model => (X -> [x=="FluxModel" ? "MLP" : x for x in X]) => :model)
transform!(df_plot, :model => (X -> [x=="LogisticRegression" ? "Linear" : x for x in X]) => :model)

ncol = length(unique(df_plot.model))
nrow = length(unique(df_plot.name))

using RCall
scale_ = 1.9
R"""
library(data.table)
df_plot <- data.table($df_plot)
model_order <- c("Linear", "MLP", "Deep Ensemble")
df_plot[,model:=factor(model, levels=model_order)]
library(ggplot2)
plt <- ggplot($df_plot) +
    geom_bar(aes(x=n, y=mean, fill=generator), stat="identity", alpha=0.5, position="dodge") +
    geom_pointrange( aes(x=n, y=mean, ymin=ymin, ymax=ymax, colour=generator), alpha=0.9, position=position_dodge(width=0.9), size=0.5) +
    facet_grid(
        rows = vars(name),
        cols =  vars(model), 
        scales = "free_y"
    ) +
    labs(y = "Value") + 
    scale_fill_discrete(name="Generator:") +
    scale_colour_discrete(name="Generator:") +
    theme(
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="bottom"
    ) +
    guides(fill=guide_legend(ncol=4))
temp_path <- file.path(tempdir(), "plot.png")
ggsave(temp_path,width=$ncol * $scale_,height=$nrow * $scale_ * 0.8) 
"""

img = Images.load(rcopy(R"temp_path"))
Images.save(joinpath(www_path,"paper_synthetic_latent_results.png"), img)
```

```{julia}
#| label: fig-mit-latent-paper
#| fig-cap: "Chart in paper"

Images.load(joinpath(www_path,"paper_synthetic_latent_results.png"))
```

### Real World

```{julia}
models = [
    :LogisticRegression, 
    :FluxModel, 
    :FluxEnsemble,
]
opt = Flux.Descent(0.01) 
generators = Dict(
    :Generic=>GenericGenerator(opt = opt, decision_threshold=0.5),
    :Latent=>REVISEGenerator(opt = opt),
    :Generic_conservative=>GenericGenerator(opt = opt, decision_threshold=0.9),
    :Gravitational=>GravitationalGenerator(opt = opt),
    :ClapROAR=>ClapROARGenerator(opt = opt)
)
```

```{julia}
max_obs = 2500
data_path = data_dir("real_world")
data_sets = load_real_world(max_obs)
choices = [
    :cal_housing, 
    :credit_default, 
    :gmsc, 
]
data_sets = filter(p -> p[1] in choices, data_sets)
```

```{julia}
using CounterfactualExplanations.DataPreprocessing: unpack
bs = 500
function data_loader(data::CounterfactualData)
    X, y = unpack(data)
    data = Flux.DataLoader((X,y),batchsize=bs)
    return data
end
model_params = (batch_norm=false,n_hidden=64,n_layers=3,dropout=true,p_dropout=0.1)
```

```{julia}
experiments = set_up_experiments(
    data_sets,models,generators; 
    pre_train_models=100, model_params=model_params, 
    data_loader=data_loader
)
```

```{julia}
n_evals = 5
n_rounds = 50
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 5
n_samples = 10000
T = 100
generative_model_params = (epochs=250, latent_dim=8)
results = run_experiments(
    experiments;
    save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, T=T, n_samples=n_samples,
    generative_model_params=generative_model_params
)
Serialization.serialize(joinpath(output_path,"results_real_world.jls"),results)
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results_real_world.jls"))
```

```{julia}
using Images
line_charts = Dict()
errorbar_charts = Dict()
for (data_name, res) in results
    plt = plot(res)
    Images.save(joinpath(www_path, "line_chart_$(data_name).png"), plt)
    line_charts[data_name] = plt
    plt = plot(res,maximum(res.output.n))
    Images.save(joinpath(www_path, "errorbar_chart_$(data_name).png"), plt)
    errorbar_charts[data_name] = plt
end
```

#### Bootstrap

```{julia}
n_bootstrap = 100
df = run_bootstrap(results, n_bootstrap; filename=joinpath(output_path,"bootstrap_real_world.csv"))
```

#### Chart in paper

@fig-mit-latent-paper shows the chart that went into the paper.

```{julia}
using DataFrames, Statistics
model_ = :FluxModel
df = DataFrame() 
for (key, val) in results
    df_ = deepcopy(val.output)
    df_.dataset .= key
    df = vcat(df,df_)
end
df = df[df.n .== maximum(df.n),:]
df = df[df.model .== model_,:]
filter!(:value => x -> !any(f -> f(x), (ismissing, isnothing, isnan)), df)
gdf = groupby(df, [:generator, :dataset, :n, :name, :scope])
df_plot = combine(gdf, :value => (x -> [(mean(x),mean(x)+std(x),mean(x)-std(x))]) => [:mean, :ymax, :ymin])
df_plot = df_plot[[name in [:mmd, :model_performance] for name in df_plot.name],:]
df_plot = df_plot[.!(df_plot.name.==:mmd .&& df_plot.scope.!=:model),:]
df_plot = mapcols(x -> typeof(x) == Vector{Symbol} ? string.(x) : x, df_plot)
transform!(df_plot, :dataset => (X -> [x=="cal_housing" ? "California Housing" : x for x in X]) => :dataset)
transform!(df_plot, :dataset => (X -> [x=="credit_default" ? "Credit Default" : x for x in X]) => :dataset)
transform!(df_plot, :dataset => (X -> [x=="gmsc" ? "GMSC" : x for x in X]) => :dataset)
transform!(df_plot, :name => (X -> [x=="mmd" ? "MMD (model)" : x for x in X]) => :name)
transform!(df_plot, :name => (X -> [x=="model_performance" ? "Performance" : x for x in X]) => :name)
transform!(df_plot, :generator => (X -> [x=="Generic" ? "Generic (γ=0.5)" : x for x in X]) => :generator)
transform!(df_plot, :generator => (X -> [x=="Generic_conservative" ? "Generic (γ=0.9)" : x for x in X]) => :generator)

ncol = length(unique(df_plot.dataset))
nrow = length(unique(df_plot.name))

using RCall
scale_ = 2.0
R"""
library(ggplot2)
plt <- ggplot($df_plot) +
    geom_bar(aes(x=n, y=mean, fill=generator), stat="identity", alpha=0.5, position="dodge") +
    geom_pointrange( aes(x=n, y=mean, ymin=ymin, ymax=ymax, colour=generator), alpha=0.9, position=position_dodge(width=0.9), size=0.5) +
    facet_grid(
        rows = vars(name),
        cols =  vars(dataset), 
        scales = "free_y"
    ) +
    labs(y = "Value") + 
    scale_fill_discrete(name="Generator:") +
    scale_colour_discrete(name="Generator:") +
    theme(
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="bottom"
    ) +
    guides(fill=guide_legend(ncol=3))
temp_path <- file.path(tempdir(), "plot.png")
ggsave(temp_path,width=$ncol * $scale_,height=$nrow * $scale_ * 0.85) 
"""

img = Images.load(rcopy(R"temp_path"))
Images.save(joinpath(www_path,"paper_real_world_results.png"), img)
```

```{julia}
#| label: fig-mit-real-paper
#| fig-cap: "Chart in paper"

Images.load(joinpath(www_path,"paper_real_world_results.png"))
```

```{julia}
generate_artifacts(output_path)
generate_artifacts(www_path)
```