### 2023-01-03 paper/sections/empirical_2.rmd

# Experiments {#empirical-2}

Below, we first present our main experimental findings regarding these questions. We conclude this section with a brief recap providing answers to all of these questions.

## Endogenous Macrodynamics

We start this section off with the key high-level observations. Across all datasets (synthetic and real), classifiers and counterfactual generators we observe either most or all of the following dynamics at varying degrees:

- Statistically significant domain and model shift as measured by MMD. 
- A deterioration in out-of-sample model performance as measured by the F-Score evaluated on a test sample. In many cases this drop in performance is substantial.
- Significant perturbations to the model parameters as well as an increase in the model's decisiveness.
- Disagreement between the original and retrained model, in some cases large.

There is also some clear heterogeneity across the results:

- The observed dynamics are generally of the highest magnitude for the linear classifier. Differences in results for the MLP and Deep Ensemble are mostly negligible. 
- The reduction in model performance appears to be most severe when classes are not perfectly separable or the initial model performance was weak, to begin with.
- Except for the Greedy generator, all other generators generally perform somewhat better overall than the baseline (Wachter) as expected.

Focusing first on synthetic data, Figure \@ref(fig:syn) presents our findings for the dataset with overlapping classes. It shows the resulting values for some of our evaluation metrics at the end of the experiment, after all $T=50$ rounds, along with error bars indicating the variation across folds.

The top row shows the estimated domain shifts. While it is difficult to interpret the exact magnitude of MMD, we can see that the values are different from zero and there is essentially no variation across our five folds. For the domain shifts, the Greedy generator induces the smallest shifts. In general, we have observed the opposite.

The second row shows the estimated model shifts, where here we have used the grid approach explained earlier. As with the domain shifts, the observed values are clearly different from zero and variation across folds is once again small. In this case, the results for this particular dataset very much reflect the broader patterns we have observed: Latent Space (LS) generators induce the smallest shifts, followed by DiCE, then Wachter and finally Greedy.

The same broad pattern also emerges in the third row: we observe the smallest deterioration in model performance for LS generators, albeit we still find a reduction in the F-Score of around 5-10 percentage points on average. Related to this, the bottom two rows indicate that the retrained classifiers disagree with their initial counterparts on the classification of up to nearly 25 per cent of the individuals. We also note that the final classifiers are more decisive, although as we noted earlier this may to some extent just be a byproduct of retraining the model throughout the experiment.

Figure \@ref(fig:syn) also indicates that the estimated effects are strongest for the simplest linear classifier, a pattern that we have observed fairly consistently. Conversely, there is virtually no difference in outcomes between the deep ensemble and the MLP. It is possible that the deep ensembles simply fail to capture predictive uncertainty well and hence counterfactual generators like Greedy, which explicitly addresses this quantity, fail to work as expected.

The findings for the other synthetic datasets are broadly consistent with the observations above. For the Moons data, the same broad patterns emerge, although in this case, the Greedy generator induces comparably strong shifts in some cases. For the Circles data, it also appears that Latent Space search yields better results. Model shifts and performance deterioration are also quantitatively smaller than what we can observe in Figure \@ref(fig:syn). For the Linearly Separable data we also find substantial domain and model shifts, but no reduction in model performance.^[You can find a granular overview of all results including bootstraps in our online companion: [https://www.paltmeyer.com/endogenous-macrodynamics-in-algorithmic-recourse/](https://www.paltmeyer.com/endogenous-macrodynamics-in-algorithmic-recourse/).]

Finally, it is also worth noting that the observed dynamics and patterns are consistent throughout the experiment. That is to say that we start observing shifts already after just a few rounds and these tend to increase proportionately for the different generators over the course of the experiment.

```{r syn, fig.cap="Results for synthetic data with overlapping classes. The shown model MMD (PP MMD) was computed over a mesh grid of 1,000 points. Error bars indicate the standard deviation across folds."}
knitr::include_graphics("www/synthetic_results.png")
```

Turning to the real-world data we will go through the findings presented in Figure \@ref(fig:real), where each column corresponds to one of the three data sets. The results shown here are for the deep ensemble, which once again largely resemble those for the MLP. Starting from the top row, we find significant domain shifts of varying magnitudes. Latent Space search induces shifts that are orders of magnitude higher than for the other generators, which generally induce significant but small shifts.

Model shifts are shown in the middle row of Figure \@ref(fig:real): the estimated PP MMD is statistically significant across the board and in some cases much larger than in others. We find no evidence that LS search helps to mitigate model shifts, as we did before for the synthetic data. Since these real-world datasets are arguably more complex than the synthetic data, the generative model can be expected to have a harder time learning the data-generating process and hence this increased difficulty appears to affect the performance of REVISE/CLUE.

The out-of-sample model performance also deteriorates across the board and substantially so: the largest average reduction in F-Scores of around 15-20 percentage points is observed for the Credit Default dataset. For this dataset we achieved the lowest initial model performance, indicating once again that weaker classifiers may be more exposed to endogenous dynamics. As with the synthetic data, the estimates for logistic regression are qualitatively in line with the above, but quantitatively even more pronounced.

```{r real, fig.cap="Results for deep ensemble using real-world datasets. The shown model MMD (PP MMD) was computed over actual samples, rather than a mesh grid. Error bars indicate the standard deviation across folds."}
knitr::include_graphics("www/real_world_results.png")
```

To recap, we answer our research questions: firstly, endogenous dynamics do emerge in our experiments (RQ \@ref(prp:shifts)) and we find them substantial enough to be considered costly (RQ \@ref(prp:costs)); secondly, the choice of the counterfactual generator matters, with Latent Space search generally having a dampening effect (RQ \@ref(prp:het)). The observed dynamics, therefore, seem to be driven by a discrepancy between counterfactual outcomes that minimize costs to individuals and outcomes that comply with the data-generating process (RQ \@ref(prp:drive)).