### 2023-01-03 paper/sections/methodology.rmd

# Gradient-Based Recourse Revisited {#method}

In this section, we first set out a generalized framework for gradient-based counterfactual search that encapsulates the various individual recourse methods we have chosen to use in our experiments (Section \@ref(method-general)). We then introduce the notion of a hidden external cost in algorithmic recourse and extend the existing framework to explicitly address this cost in the counterfactual search objective (Section \@ref(method-collective)).

## From individual recourse ... {#method-general}

We have chosen to focus on gradient-based counterfactual search for two reasons: firstly, they can be seen as direct descendants of our baseline method (Wachter); secondly, gradient-based search is particularly well-suited for differentiable black-box models like deep neural networks, which we focus on in this work. In particular, we include the following generators in our simulation experiments below: **REVISE** [@joshi2019realistic], **CLUE** [@antoran2020getting], **DiCE** [@mothilal2020explaining] and a greedy approach that relies on probabilistic models [@schut2021generating]. Our motivation for including these different generators in our analysis is that they all offer slightly different approaches to generating meaningful counterfactuals for differentiable black-box models. We hypothesize that generating more **meaningful** counterfactuals should mitigate the endogenous dynamics illustrated in Figure \@ref(fig:poc) in Section \@ref(intro). This intuition stems from the underlying idea that more meaningful counterfactuals are generated by the same or at least a very similar data-generating process as the training data. All else equal, counterfactuals that fulfil this basic requirement should be less prone to trigger shifts.

As we will see next, all of them can be described by the following generalized form of Equation \@ref(eq:general):

\begin{equation}
\begin{aligned}
\mathbf{s}^\prime &= \arg \min_{\mathbf{s}^\prime \in \mathcal{S}} \left\{  {\text{yloss}(M(f(\mathbf{s}^\prime)),y^*)}+ \lambda {\text{cost}(f(\mathbf{s}^\prime)) }  \right\} (\#eq:general)
\end{aligned} 
\end{equation}

Here $\mathbf{s}^\prime=\left\{s_k^\prime\right\}_K$ is a $K$-dimensional array of counterfactual states and $f: \mathcal{S} \mapsto \mathcal{X}$ maps from the counterfactual state space to the feature space. In Wachter, the state space is the feature space: $f$ is the identity function and the number of counterfactuals $K$ is one. Both REVISE and CLUE search counterfactuals in some latent space $\mathcal{S}$ instead of the feature space directly. The latent embedding is learned by a separate generative model that is tasked with learning the data-generating process (DGP) of $X$. In this case, $f$ in Equation \@ref(eq:general) corresponds to the decoder part of the generative model, that is the function that maps back from the latent space to inputs. Provided the generative model is well-specified, traversing the latent embedding typically yields meaningful counterfactuals since they are implicitly generated by the (learned) DGP [@joshi2019realistic].

CLUE distinguishes itself from REVISE and other counterfactual generators in that it aims to minimize the predictive uncertainty of the model in question, $M$. To quantify predictive uncertainty, Antoran et al. [@antoran2020getting] rely on entropy estimates for probabilistic models. The greedy approach proposed by Schut et al. @schut2021generating, which we refer to as **Greedy**, also works with the subclass of models $\tilde{\mathcal{M}}\subset\mathcal{M}$ that can produce predictive uncertainty estimates. The authors show that in this setting the cost function $\text{cost}(\cdot)$ in Equation \@ref(eq:general) is redundant and meaningful counterfactuals can be generated in a fast and efficient manner through a modified Jacobian-based Saliency Map Attack (JSMA). Schut et al. [@schut2021generating] also show that by maximizing the predicted probability of $x^\prime$ being assigned to target class $y^*$, we also implicitly minimize predictive entropy (as in CLUE). In that sense, CLUE can be seen as equivalent to REVISE in the Bayesian context and we shall therefore refer to both approaches collectively as **Latent Space** generators^[In fact, there are several other recently proposed approaches to counterfactual search that also broadly fall in this same category. They largely differ with respect to the chosen generative model: for example, the generator proposed by Dombrowski et al. @dombrowski2021diffeomorphic relies on normalizing flows.].

Finally, DiCE [@mothilal2020explaining] distinguishes itself from all other generators considered here in that it aims to generate a diverse set of $K>1$ counterfactuals. Wachter et al. [@wachter2017counterfactual] show that diverse outcomes can in principle be achieved simply by rerunning counterfactual search multiple times using stochastic gradient descent (or by randomly initializing the counterfactual)^[Note that \@ref(eq:general) naturally lends itself to that idea: setting $K$ to some value greater than one and using the Wachter objective essentially boils down to computing multiple counterfactuals in parallel. Here, $yloss(\cdot)$ is first broadcasted over elements of $\mathbf{s}^\prime$ and then aggregated. This is exactly how counterfactual search is implemented in [`CounterfactualExplanations.jl`](https://github.com/pat-alt/CounterfactualExplanations.jl).]. In @mothilal2020explaining diversity is explicitly proxied via Determinantal Point Processes (DDP): the authors simply introduce DDP as a component of the cost function $\text{cost}(\mathbf{s}^\prime)$ and thereby produce counterfactuals $s_1, ..., s_K$ that look as different from each other as possible. The implementation of DiCE in our library of choice---[`CounterfactualExplanations.jl`](https://github.com/pat-alt/CounterfactualExplanations.jl)---uses that exact approach. It is worth noting that for $k=1$, DiCE reduces to Wachter since the DDP is constant and therefore does not affect the objective function in Equation \@ref(eq:general).

## ... towards collective recourse {#method-collective}

All of the different approaches introduced above tackle the problem of Algorithmic Recourse from the perspective of one single individual^[DiCE recognizes that different individuals may have different objective functions, but it does not address the interdependencies between different individuals.]. To explicitly address the issue that individual recourse may affect the outcome and prospect of other individuals, we propose to extend Equation \@ref(eq:general) as follows:

\begin{equation}
\begin{aligned}
\mathbf{s}^\prime &= \arg \min_{\mathbf{s}^\prime \in \mathcal{S}} \{ {\text{yloss}(M(f(\mathbf{s}^\prime)),y^*)} \\ &+ \lambda_1 {\text{cost}(f(\mathbf{s}^\prime))} + \lambda_2 {\text{extcost}(f(\mathbf{s}^\prime))} \}  (\#eq:collective)
\end{aligned} 
\end{equation}

Here $\text{cost}(f(\mathbf{s}^\prime))$ denotes the proxy for private costs faced by the individual as before and $\lambda_1$ governs to what extent that private cost ought to be penalized. The newly introduced term $\text{extcost}(f(\mathbf{s}^\prime))$ is meant to capture and address external costs incurred by the collective of individuals in response to changes in $\mathbf{s}^\prime$. The underlying concept of private and external costs is borrowed from Economics and well-established in that field: when the decisions or actions by some individual market participant generate external costs, then the market is said to suffer from negative externalities and is considered inefficient [@pindyck2014microeconomics]. We think that this concept describes the endogenous dynamics of algorithmic recourse observed here very well. As with individual recourse, the exact choice of $\text{extcost}(\cdot)$ is not obvious, nor do we intend to provide a definite answer in this work, if such even exists. That being said, we do propose a few potential mitigation strategies in Section \@ref(mitigate).