### 2023-02-02 README.qmd
updated README

---
format: commonmark
bibliography: bib.bib
---

# Endogenous Macrodynamics in Algorithmic Recourse

This repository contains all code, notebooks, data and empirical results for our conference paper "Endogenous Macrodynamics in Algorithmic Recourse" [@altmeyer2023endogenous].

Below is a list of relevant resources hosted in this repository:

1. [Paper](paper/paper.pdf) in this repo 
2. [Paper](https://openreview.net/pdf?id=-LFT2YicI9v) on OpenReview
3. [Online Companion](https://www.paltmeyer.com/endogenous-macrodynamics-in-algorithmic-recourse/)
4. [IEEE SaTML Presentation Slides](https://www.paltmeyer.com/content/talks/posts/2023-ieee-satml/presentation.html)
5. [IEEE SaTML Poster](dev/poster/poster.pdf)
6. Software: [`AlgorithmicRecourseDynamics.jl`](https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl) and [`CounterfactualExplanations.jl`](https://github.com/pat-alt/CounterfactualExplanations.jl)

## Motivation

The chart below illustrates what we define as macrodynamics in Algorithmic Recourse: (a) we have a simple linear classifier trained for binary classification where samples from the negative class ($y=0$) are marked in orange and samples of the positive class ($y=1$) are marked in blue; (b) the implementation of AR for a random subset of individuals leads to a noticeable domain shift; (c) as the classifier is retrained we observe a corresponding model shift; (d) as this process is repeated, the decision boundary moves away from the target class.

![](paper/www/poc.png)

## Abstract

Existing work on Counterfactual Explanations (CE) and Algorithmic Recourse (AR) has largely focused on single individuals in a static environment: given some estimated model, the goal is to find valid counterfactuals for an individual instance that fulfill various desiderata. The ability of such counterfactuals to handle dynamics like data and model drift remains a largely unexplored research challenge. There has also been surprisingly little work on the related question of how the actual implementation of recourse by one individual may affect other individuals. Through this work, we aim to close that gap. We first show that many of the existing methodologies can be collectively described by a generalized framework. We then argue that the existing framework does not account for a hidden external cost of recourse, that only reveals itself when studying the endogenous dynamics of recourse at the group level. Through simulation experiments involving various state-of-the-art counterfactual generators and several benchmark datasets, we generate large numbers of counterfactuals and study the resulting domain and model shifts. We find that the induced shifts are substantial enough to likely impede the applicability of Algorithmic Recourse in some situations. Fortunately, we find various strategies to mitigate these concerns. Our simulation framework for studying recourse dynamics is fast and open-sourced.

## Key Findings

- Our findings indicate that state-of-the-art approaches to Algorithmic Recourse induce substantial domain and model shifts.
- We would argue that the expected external costs of individual recourse should be shared by all stakeholders.
- A straightforward way to achieve this is to penalize external costs in the counterfactual search objective function (Equation 4).
- Various simple strategies based on this notion can be effectively used to mitigate shifts.

## Proposed Mitigation Strategies

By introducing a second penalty term in the counterfactual search objective, we can explicitly penalize external costs. The figure below illustrates how the mitigation strategies compared to the baseline approach, that is, Wachter (Generic) with γ = 0.5: choosing a higher decision threshold pushes the counterfactual a little further into the target domain; this effect is even stronger for ClaPROAR; finally, using the Gravitational generator the counterfactual ends up all the way inside the target domain. Find out more in the [paper](paper/paper.pdf).

![](paper/www/mitigation.png)

## References